---
title: "Clustering with Geolocation Saturation in the AFH Market"
author: "<a href='https://evanhatt.github.io' target='_blank'>Ellie Van Hattem Barasa</a> x <a href='' target='_blank'>Revival, co.</a>"
date: "10/1/2024"
mail: "ellievanhattembarasa@gmail.com"
#linkedin: ""
#twitter: ""
github: "evanhatt"
home: "https://github.com/evanhatt"
logo: "images/combined_logo.png"
output:
  epuRate::epurate:
    #css: style.css
    toc: TRUE
    number_sections: FALSE
    code_folding: "hide"
    #includes: 
        #before_body: header.html
---

# Welcome

The purpose of this analysis is to look at the clustering of current Adult Family Homes for future market potentional in Benton County.  
This Analysis uses R vizualization and Python scripting together, utilizing the reticulate package. The data was filitered and cleaned using Python. A Python function was also created for the clustering algorithm. Both files can be found in my [GitHub Repository for this analysis.](https://github.com/evanhatt/AFH_WA)


Disclaimer  
I do not own the rights to this data. I utilized government collected data on current open and active AFH homes in Washington. 

## A little about who Revival AFH is

Revival Adult Family Homes is a business that puts its residents first in every aspect with a huge goal: To provide an exceptional standard of care by treating residents like their own senior family members. They built a system of care that aims to elevate the standard of care and living through live-in care givers, organic and freshly prepared meals, and specialized attention to each resident. And they want to continue helping the golden year community with more rooms!  They are also currently in the development phase for an app that allows CNA's to talk and negotiate directly with in need Facilities. Be on the look out for their launch!

## About AFH Market in WA

Revival AFH LLC is looking to open another home in Benton County, Washington. After establishing their first home in Kennewick, they reached out for insights on the current market saturation of AFH (Adult Family Homes) in the area and geographical advice for their next business expansion plans.


```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

####### ALL LIBRARIES USED ###########
###### DATA MANIPULATION ######
options(stringsAsFactors = FALSE)
Library(Pandas) # for main data manipulations 
Library(Geopandas) # for spatial and geographic data
Library(Matplotlib.pyplot) # for creating static, animated, and interactive visualizations
source_python("euclidean_script.py") # python script

###### GRAPHING ######
#packages 
Library(Pandas) # for main data manipulations  
Library(Seaborn) # for colors in graphs
Library(Folium) # For plot markers and geographic data
Library(Contextily)  # For adding a basemap
Library(Plotly.express) # Data binding, Axes labels, Colors, and Legends.

######################################

# titles
latitude_col = 'Latitude'           # Replace with actual latitude column name
longitude_col = 'Longitude'         # Replace with actual longitude column name
location_county_col = 'Location county'  # Replace with actual county column name
location_address_col = 'Location address'  # Column for addresses
facility_type_col = 'Facility type'  # Column for facility type

# Filter the dataset for Benton County and Facility type 'AF'
benton_county_afh = raw[(raw[location_county_col] == 'Benton') & (raw[facility_type_col] == 'AF')]


# Convert the dataframe into a GeoDataFrame
gdf = gpd.GeoDataFrame(
    benton_county_afh,
    geometry=gpd.points_from_xy(benton_county_afh[longitude_col], benton_county_afh[latitude_col])
)

# Set the coordinate reference system to WGS84 (latitude/longitude)
gdf.set_crs(epsg=4326, inplace=True)

# Plot the Benton County AFHs on a map
fig, ax = plt.subplots(figsize=(10, 10))

# Plot the AFH locations
gdf.plot(ax=ax, color='blue', marker='o', markersize=50, alpha=0.7, label='AFH Locations')

# Add a basemap (OpenStreetMap tiles)
ctx.add_basemap(ax, crs=gdf.crs.to_string(), source=ctx.providers.OpenStreetMap.Mapnik)

# Customize the plot
plt.title('Adult Family Home Locations in Benton County')
plt.xlabel('Longitude')
plt.ylabel('Latitude')

# Show the plot
plt.legend()
plt.tight_layout()
plt.show()
```

# How many homes are in Richland and Kennewick?

Lets start with getting a baseline for the volume of litter collected on each of our four days.

---
title: "AFH Facility Count in ZIP Code 99352"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Load necessary libraries
library(dplyr)
library(knitr)

# Assuming 'raw' is your dataset containing the AFH data

# Filter the data for AFH facilities with ZIP code 99352 and Facility type 'AF'
afh_zip_99352 <- raw %>%
  filter(`Mail ZIP Code` == "99352", `Facility type` == "AF")

# Count the number of AFH facilities in the 99352 ZIP code
num_afh_zip_99352 <- nrow(afh_zip_99352)

r num_afh_zip_99352
```
---
title: "AFH Facility Count in ZIP Code 99337"
author: "Ellie Barasa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Load necessary libraries
library(dplyr)
library(knitr)

# Assuming 'raw' is your dataset containing the AFH data

# Filter the data for AFH facilities with ZIP code 99337 and Facility type 'AF'
afh_zip_99337 <- raw %>%
  filter(`Mail ZIP Code` == "99337", `Facility type` == "AF")

# Count the number of AFH facilities in the 99337 ZIP code
num_afh_zip_99337 <- nrow(afh_zip_99337)

r num_afh_zip_99337
```


Provide Specific analysis conlcusion based on the data (explain it) 



# What about Geolocation objects?

Geolocation Objects are defined to be hierarchically grouped by County, Zipcode, and City. 
For our clustering algorithm, we will filter for Benton County and use zipcodes 99352 for Richland and 99337 for Kennweick for difference assesments.


From here, we will cycle through each tracked geolocation and discorver AFH density in each City in relation to each other, via straight line distance with an adjustment for longitude and latitude. Each AFH will then be considered part of their city or zipcode cluster.

[This script was created in Python and saved as euclidean_script.py](https://github.com/evanhatt/AFH_WA/blob/master/euclidean_script.py), which can be found in my [GitHub repository for this analysis](https://github.com/evanhatt/AFH_WA). The Python script is being used directly in R as a callable function using the Reticulate package! We could have also created the function directly in the R Markdown file; however, it is much easier to just import it directly as a function. So lets cluster our AFHs!


```{r}
# clustering data points with python script
clustered_data <- raw %>% 
  subset(city == 'Redwood City') %>% # this is a different area
  get_euclidean(collection, litter) # python function

#########################
import matplotlib.pyplot as plt

# Filter the data for Benton County and Facility type 'AF'
benton_afh = raw[(raw['Location county'] == 'Benton') & (raw['Facility type'] == 'AF')]

# Group the data by 'Mail ZIP Code' and count the number of facilities in each ZIP code
zip_counts = benton_afh.groupby('Mail ZIP Code').size().reset_index(name='AFH count')

# Create a horizontal lollipop chart
plt.figure(figsize=(10, 6))

# Plot horizontal lines (the lollipop stick)
plt.hlines(y=zip_counts['Mail ZIP Code'], xmin=0, xmax=zip_counts['AFH count'], color='blue', linewidth=2)

# Plot circles (the lollipop head)
plt.scatter(zip_counts['AFH count'], zip_counts['Mail ZIP Code'], color='red', s=100)

# Add labels and title
plt.ylabel('ZIP Code')
plt.xlabel('Number of AFHs')
plt.title('Number of AFHs in Benton County Grouped by ZIP Code')

# Show the plot
plt.tight_layout()
plt.show()

```
 



# What does this all mean? 

In summary, 


## My recomendations for next location for an AFH for Revival Homes 


Finally, 


## A fun Sunburst graph!

Finally, I will leave you with this fun sunburst graph to play with. You can hover over each section to further subset........ Have fun!
```{r}
# getting data for sunburst plot
sunburst <- clustered_data %>% 
  
sunburst
```

-------------------------

-------------------------

-------------------------

-------------------------

-------------------------

Below are the code chunks for the makings and simple statistics used throughout this analysis. They are shoved to the bottom, so not to muddy up the written portion.


## The head of the raw data


```{r}
raw %>% head()
```


##  Total number of AFH in Benton County


```{r}


###############
# 
nDistance %>% nrow()
```


## 

```{r}
###############
# 

```


##

```{r}
############
# 
```


## 


```{r}

```



## 


```{r}

# 




# 

# 


```
